Bayesian model comparison for rare variant association studies of multiple phenotypes

### 2018-01-29 tex/manuscriptbiorxiv.tex
Updates for comments from MP

\documentclass[10pt]{article}
%\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

@@

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
%\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
%\fancyheadoffset[L]{2.25in}
%\fancyfootoffset[L]{2.25in}
%\lfoot{\sf PLOS}

@@

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/Figure_02.pdf}
\caption{{\bf Simulation studies.}
A: Comparison of $-\log_{10}$(p-values) from frequentist $\textrm{BF}_{\textrm{MRP}}$ approximation for an independent effects and a similar effects model to commonly used gene-based statistical tests (skat and burden). B: Comparison of log10(Bayes Factors) obtained when raw genotype and phenotype data is available to a scenario where summary statistics only was available and similar effects across studies is assumed. C: From single variant and single phenotype to multiple variants and multiple phenotypes gene discovery: ROC curves for detecting gene association to any of the phenotypes using single variant/single phenotype association (blue) to multiple variants and multiple phenotypes association (purple). D: ROC curves for detecting gene association when incorporating prior mean of genetic effects (orange) to identify protective alleles.}
\label{simresults}
\end{figure}

@@

\subsubsection*{Genetic Correlations}
We calculated the genetic between the two groups of traits (asthma, eosinophil counts, FVC, FEV$_1$ and glaucoma, intra-ocular pressure, corneal resistance factor) using the MultiVariate Polygenic Mixture Model (MVPMM)~\cite{DeBoever2017}. Briefly, MVPMM estimates genetic correlation given GWAS summary statistics (effect size and standard error of effect size estimate) by modeling GWAS summary statistics as generated from one of two mixture components. Summary statistics from variants in the null component are modeled as being drawn from a multivariate normal distribution with zero mean and covariance matrix that captures correlation in the summary statistics due to the use of shared subjects or other sources of correlation. Summary statistics from variants in the non-null component are modeled as being drawn from a multivariate normal distribution with zero mean, but the covariance matrix for the non-null component combines the covariance matrix from the null component with another covariance matrix that captures the genetic correlation between the phenotypes being considered.

\subsubsection*{UK Biobank Asthma and Glaucoma Applications}
We applied MRP for both sets of traits assuming both an independent effects model and a similar effects model. We applied Bayesian model averaging to these results by averaging the BF from each model for each gene; these results are reported in the main text. The results for each individual model are included in the Supporting Information. For the Manhattan plots and tables, we removed any genes with non-unique gene symbols. In cases where genes overlapped such that they shared rare variants and therefore the same BF, we removed one gene.

@@

%% FIGURE: ASTHMA MANHATTAN AND FOREST FOR BMA
\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/Figure_03.pdf}
\caption{{\bf Results for asthma application.}
$\log_{10}$ Bayes Factors from applying MRP and Bayesian model averaging to summary statistics for missense and protein-truncating variants from (A) asthma (HC382), (B) eosinophil counts (INI30150), (C) forced vital capacity (FVC, INI3062), (D) forced expiratory volume in 1-second (FEV$_1$, INI3063), (E) all four traits jointly, and (F) all four traits jointly with focus on protective effects. The four genes outside of chromosome 6 with the largest Bayes Factors greater than three are labeled in each plot. Only $\log_{10}$ Bayes Factors greater than -5 are plotted. (F) $-\log_{10}$ $p$-values (left panel) and estimated effect sizes with 95\% confidence intervals (right panel) for missense variants and PTVs in \textit{IL33} for each phenotype}
\label{asthma_bma}
\end{figure}

\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{asthma_bma}.}}
\begin{tabular}{|l+l|l|l|l|l|l|}
\hline
{\bf Gene} & {\bf Joint, protective} & {\bf Joint} & {\bf Eosinophil count} & {\bf FVC} & {\bf FEV$_1$} & {\bf Asthma} \\
\thickhline
\textit{IL33}         &              29.4 &   29.3 &             30.6 &  -2.3 &  -2.2 &    8.1 \\
\hline
\textit{CCR3}         &               3.1 &    3.3 &              7.4 &  -1.4 &  -1.6 &   -0.5 \\
\hline
\textit{RP11-39K24.9} &               0.8 &    1.8 &              4.9 &  -0.1 &  -0.4 &    0.3 \\
\hline
\textit{SCMH1}        &               0.5 &    4.7 &             -1.5 &   7.7 &   3.8 &   -0.7 \\
\hline
\textit{MUSTN1}       &               0.4 &    1.1 &             -1.2 &   2.9 &   2.9 &   -0.6 \\
\hline
\textit{ZFAT}         &               0.3 &    1.3 &             -2.0 &   4.7 &   3.1 &   -0.4 \\
\hline
\textit{ELN}          &               0.2 &    2.5 &             -1.0 &   5.6 &   2.9 &   -0.6 \\
\hline
\textit{C14orf39}     &              -0.7 &   -0.0 &             -1.1 &   3.5 &   2.5 &    0.0 \\
\hline
\textit{TMEM110}      &              -0.9 &    1.1 &             -1.0 &   3.3 &   3.1 &   -0.6 \\
\hline
\textit{IL17RA}       &              -4.4 &    3.1 &              7.9 &  -2.7 &  -2.5 &   -1.1 \\
\hline
\textit{IL18RAP}      &              -9.6 &   -0.9 &             -1.0 &  -1.6 &  -1.7 &    3.5 \\
\hline
\textit{ATP2A3}       &             -11.9 &   -0.8 &             -1.2 &  -2.1 &  -2.2 &    5.8 \\
\hline
\textit{FLG}          &             -20.1 &  -17.2 &             -6.0 &  -7.4 &  -8.5 &    4.0 \\
\hline
\end{tabular}
\label{asthma_table}
\end{adjustwidth}
\end{table}

\subsubsection*{Glaucoma, intra-ocular pressure, and corneal resistance factor}
We also applied MRP to missense variants and PTVs for glaucoma, intra-ocular pressure, and corneal resistance factor as well as performing joint analyses. Elevated intra-ocular pressure is a measure of the fluid pressure in the eye. Elevated intra-ocular pressure is associated with glaucoma risk and has been linked to genetic variants associated with glaucoma \cite{doi:10.1001/jama.2014.3192}. Corneal resistance factor is a measure of the cornea's ability to resist mechanical stress and has been associated with glaucoma presence and severity \cite{CXO:CXO414, MANSOURI2012419, GriseDulac:2012bo}. While the individual glaucoma analysis did not yield any associations with $\log_{10}$ BF greater than three, the joint analysis identified rare coding variants in both \textit{ANGPTL7}, \textit{WNT10A}, and \textit{KLHL22} as associated with glaucoma (Fig~\ref{glaucoma_bma}A-D). Applying the MRP model to identify protective effects also identified \textit{ANGPTL7} and \textit{WNT10A} but not \textit{KLHL22} (Fig~\ref{glaucoma_bma}E). Expression of \textit{ANGPTL7} is upregulated in glaucome and has been proposed to regulate intra-ocular pressure and glaucoma risk \cite{Comes:2011ex, Kuchtey:2008ek}. \textit{KLHL22} has not been previously associated with glaucoma though a suggestive association was reported for retinopathy in individuals without diabetes \cite{10.1371/journal.pone.0054232}. The GWAS summary statistics for the rare variants in \textit{ANGPTL7} suggest that two missense variants, rs28991002 and rs28991009, appear to drive the association with glaucoma (Fig~\ref{glaucoma_bma}F). These results illustrate that applying model comparison to multiple traits simultaneously can prioritize the discovery of protective modifiers of disease.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/Figure_04.pdf}
\caption{{\bf Results for glaucoma application.}
$\log_{10}$ Bayes Factors from applying MRP and Bayesian model averaging to summary statistics for missense and protein-truncating variants from (A) glaucoma (HC276), (B) intra-ocular pressure (INI5263), (C) corneal resistance factor (INI5265), and (D) all three traits jointly. (E) shows the results of a joint analysis focused on finding rare variants that protect against glaucoma. The genes outside of chromosome 6 with with Bayes Factor greater than three are indicated by arrows. Only $\log_{10}$ Bayes Factors greater than zero are plotted. F: $-\log_{10}$ $p$-values (left panel) and estimated effect sizes with 95\% confidence intervals (right panel) for missense variants and PTVs in \textit{ANGPTL7} for all three phenotypes.}
\label{glaucoma_bma}
\end{figure}

% %% TABLE: GLAUCOMA GENES OF INTEREST
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{glaucoma_bma}.}}
\begin{tabular}{|l+p{3cm}|p{1cm}|p{1.6cm}|p{3.8cm}|p{3.4cm}|}
\hline
{\bf Gene} & {\bf Joint, protective} & {\bf Joint} & {\bf Glaucoma} & {\bf Intra-ocular pressure, Goldmann-correlated} & {\bf Corneal resistance factor} \\
\thickhline
\textit{ANGPTL7} &              10.1 &   9.4 &      1.7 &                                       10.1 &                       6.2 \\
\hline
\textit{KLHL22}  &               3.1 &   3.6 &     -0.2 &                                        3.9 &                       2.2 \\
\hline
\textit{GLI3}    &               2.9 &   2.9 &     -0.3 &                                       -0.7 &                       3.2 \\
\hline
\textit{WNT10A}  &               2.2 &   3.0 &     -0.3 &                                       -0.5 &                       3.5 \\
\hline
\textit{RNASE2}  &              -1.3 &  -0.3 &      0.4 &                                       -1.0 &                      -1.1 \\
\hline
\textit{FUK}     &              -6.5 &  -2.3 &      2.5 &                                       -2.1 &                      -2.1 \\
\hline
\end{tabular}
\label{table2}
\end{adjustwidth}
\end{table}

@@

\begin{figure}[H]
\includegraphics[width=\textwidth]{../figures/Figure_S01.pdf}
\end{figure}
\paragraph*{S1 Fig.}
\label{hdf5_fig}
{\bf HDF5 Implementation.} Our HDF5 implementation contained the following components: first, a group with one table per annotation file. All effect size (beta) values and study-specific annotations were contained here, and the number of tables is limited by $S\ \textrm{(the number of studies)} \times K\ \textrm{(the number of traits)}$. Second, a group with site-site covariance data. While these covariance matrices may have dimension $M\ \textrm{(the number of variants)} \times M$, we store the data as tables, each row specifiying the covariance between two variants. The number of tables should be the same as the previous set, capped by $S\ \textrm{(the number of studies)} \times K\ \textrm{(the number of traits)}$. Third, we store one table with sigma values for each study/phenotype combination. In the event that the traits were rank-normal transformation was performed these sigma values are equal to 1. These are used to compute correlation between two datasets. Finally, we store a matrix/table pair for Vy null and its index. The Vy null matrix has dimensions $ (S \times K) \times (S \times K)$ each entry specifying the estimated correlation of effect sizes between two datasets. The index table encodes row/column position of each dataset.

%% FIGURE: ASTHMA MANHATTAN AND FOREST FOR INDEPENDENT EFFECTS
\begin{figure}[H]
\includegraphics[width=\textwidth]{../figures/Figure_S02.pdf}
\end{figure}
\paragraph*{S2 Fig.}
\label{asthma_independent}
{\bf Results for independent effects model applied to asthma, eosinophil counts, FEV$_1$, and FVC.} $\log_{10}$ Bayes Factors from applying MRP independent effects model to summary statistics for missense and protein-truncating variants from (A) asthma (HC382), (B) eosinophil counts (INI30150), (C) forced vital capacity (FVC, INI3062), (D) forced expiratory volume in 1-second (FEV$_1$, INI3063), (E) all four traits jointly, and (F) all four traits jointly with focus on protective effects. The four genes outside of chromosome 6 with the largest Bayes Factors greater than three are labeled in each plot. Only $\log_{10}$ Bayes Factors greater than -5 are plotted.

%% FIGURE: ASTHMA MANHATTAN AND FOREST FOR SIMILAR EFFECTS
\begin{figure}[H]
\includegraphics[width=\textwidth]{../figures/Figure_S03.pdf}
\end{figure}
\paragraph*{S3 Fig.}
\label{asthma_similar}
{\bf Results for similar effects model applied to asthma, eosinophil counts, FEV$_1$, and FVC.} $\log_{10}$ Bayes Factors from applying MRP similar effects model to summary statistics for missense and protein-truncating variants from (A) asthma (HC382), (B) eosinophil counts (INI30150), (C) forced vital capacity (FVC, INI3062), (D) forced expiratory volume in 1-second (FEV$_1$, INI3063), (E) all four traits jointly, and (F) all four traits jointly with focus on protective effects. The four genes outside of chromosome 6 with the largest Bayes Factors greater than three are labeled in each plot. Only $\log_{10}$ Bayes Factors greater than -5 are plotted.

\begin{figure}[H]
\includegraphics[width=\textwidth]{../figures/Figure_S04.pdf}
\end{figure}
\paragraph*{S4 Fig.}
\label{glaucoma_similar}
{\bf Results for independent effects model applied to glaucoma intra-ocular pressure, and corneal resistance factor.} $\log_{10}$ Bayes Factors from applying MRP similar effects model to summary statistics for missense and protein-truncating variants from (A) glaucoma (HC276), (B) intra-ocular pressure (INI5263), (C) corneal resistance factor (INI5265), and (D) all three traits jointly. (E) shows the results of a joint analysis focused on finding rare variants that protect against glaucoma. The genes outside of chromosome 6 with with Bayes Factor greater than three are indicated by arrows. Only $\log_{10}$ Bayes Factors greater than zero are plotted.

\begin{figure}[H]
\includegraphics[width=\textwidth]{../figures/Figure_S05.pdf}
\end{figure}
\paragraph*{S5 Fig.}
\label{glaucoma_independent}
{\bf Results for similar effects model applied to glaucoma intra-ocular pressure, and corneal resistance factor.} $\log_{10}$ Bayes Factors from applying MRP independent effects model to summary statistics for missense and protein-truncating variants from (A) glaucoma (HC276), (B) intra-ocular pressure (INI5263), (C) corneal resistance factor (INI5265), and (D) all three traits jointly. (E) shows the results of a joint analysis focused on finding rare variants that protect against glaucoma. The genes outside of chromosome 6 with with Bayes Factor greater than three are indicated by arrows. Only $\log_{10}$ Bayes Factors greater than zero are plotted.

>>>>

\documentclass[10pt,letterpaper]{article}
%\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

@@

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
%\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
%\fancyheadoffset[L]{2.25in}
%\fancyfootoffset[L]{2.25in}
%\lfoot{\sf PLOS}

@@

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/final/Figure_02.pdf}
\caption{{\bf Simulation studies.}
A: Comparison of $-\log_{10}$(p-values) from frequentist $\textrm{BF}_{\textrm{MRP}}$ approximation for an independent effects and a similar effects model to commonly used gene-based statistical tests (skat and burden). B: Comparison of log10(Bayes Factors) obtained when raw genotype and phenotype data is available to a scenario where summary statistics only was available and similar effects across studies is assumed. C: From single variant and single phenotype to multiple variants and multiple phenotypes gene discovery: ROC curves for detecting gene association to any of the phenotypes using single variant/single phenotype association (green) to multiple variants and multiple phenotypes association (purple). D: ROC curves for detecting gene association when incorporating prior mean of genetic effects (orange) to identify protective alleles.}
\label{simresults}
\end{figure}

@@

\subsubsection*{Genetic Correlations}
We calculated the genetic correlation between the two groups of traits (asthma, eosinophil counts, FVC, FEV$_1$ and glaucoma, intra-ocular pressure, corneal resistance factor) using the MultiVariate Polygenic Mixture Model (MVPMM)~\cite{DeBoever2017}. Briefly, MVPMM estimates genetic correlation given GWAS summary statistics (effect size and standard error of effect size estimate) by modeling GWAS summary statistics as generated from one of two mixture components. Summary statistics from variants in the null component are modeled as being drawn from a multivariate normal distribution with zero mean and covariance matrix that captures correlation in the summary statistics due to the use of shared subjects or other sources of correlation. Summary statistics from variants in the non-null component are modeled as being drawn from a multivariate normal distribution with zero mean, but the covariance matrix for the non-null component combines the covariance matrix from the null component with another covariance matrix that captures the genetic correlation between the phenotypes being considered.

\subsection*{UK Biobank Asthma and Glaucoma Applications}
We applied MRP for both sets of traits assuming both an independent effects model and a similar effects model. We applied Bayesian model averaging to these results by averaging the BF from each model for each gene; these results are reported in the main text. The results for each individual model are included in the Supporting Information. For the Manhattan plots and tables, we removed any genes with non-unique gene symbols. In cases where genes overlapped such that they shared rare variants and therefore the same BF, we removed one gene.

\subsubsection*{Variant Filtering}
We used the $\texttt{variant\_filter\_table.tsv}$ file available at https://github.com/rivas-lab/public-resources (6f9f726) to filter variants on the UK Biobank array for use with MRP. We first chose variants with minor allele frequency less than $1\%$. We then filtered out all variants with $\texttt{all\_filters}$ less than one. This removes variants with missingness greater than $1\%$ (calculated on an array-specific basis for array-specific variants) or Hardy-Weinberg equilibrium $p < 10^{-7}$. This also removes some PTVs for which manual inspection revealed irregular cluster plots \cite{DeBoever179762}. We LD pruned the variants by only using variants with $\texttt{ld}$ equal to one. We included missense variants and PTVs indicated by the following annotations: $\texttt{missense\_variant}$, 
$\texttt{stop\_gained}$, $\texttt{frameshift\_variant}$, $\texttt{splice\_acceptor\_variant}$, $\texttt{splice\_donor\_variant}$, $\texttt{splice\_region\_variant}$, $\texttt{start\_lost}$, $\texttt{stop\_lost}$.

@@

%% FIGURE: ASTHMA MANHATTAN AND FOREST FOR BMA
\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/final/Figure_03.pdf}
\caption{{\bf Results for asthma application.}
$\log_{10}$ Bayes Factors from applying MRP and Bayesian model averaging to summary statistics for missense and protein-truncating variants from (A) asthma (HC382), (B) eosinophil counts (INI30150), (C) forced vital capacity (FVC, INI3062), (D) forced expiratory volume in 1-second (FEV$_1$, INI3063), (E) all four traits jointly, and (F) all four traits jointly with focus on protective effects. The four genes outside of chromosome 6 with the largest Bayes Factors greater than three are labeled in each plot. Only $\log_{10}$ Bayes Factors greater than -5 are plotted. (F) $-\log_{10}$ $p$-values (left panel) and estimated effect sizes with 95\% confidence intervals (right panel) for missense variants and PTVs in \textit{IL33} for each phenotype}
\label{asthma_bma}
\end{figure}

\begin{table}[!ht]
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{asthma_bma}.}}
\begin{tabular}{|l+l|l|l|l|l|l|}
\hline
{\bf Gene} & {\bf Joint, protective} & {\bf Joint} & {\bf Eosinophil count} & {\bf FVC} & {\bf FEV$_1$} & {\bf Asthma} \\
\thickhline
\textit{IL33}         &              29.4 &   29.3 &             30.6 &  -2.3 &  -2.2 &    8.1 \\
\hline
\textit{CCR3}         &               3.1 &    3.3 &              7.4 &  -1.4 &  -1.6 &   -0.5 \\
\hline
\textit{RP11-39K24.9} &               0.8 &    1.8 &              4.9 &  -0.1 &  -0.4 &    0.3 \\
\hline
\textit{SCMH1}        &               0.5 &    4.7 &             -1.5 &   7.7 &   3.8 &   -0.7 \\
\hline
\textit{MUSTN1}       &               0.4 &    1.1 &             -1.2 &   2.9 &   2.9 &   -0.6 \\
\hline
\textit{ZFAT}         &               0.3 &    1.3 &             -2.0 &   4.7 &   3.1 &   -0.4 \\
\hline
\textit{ELN}          &               0.2 &    2.5 &             -1.0 &   5.6 &   2.9 &   -0.6 \\
\hline
\textit{C14orf39}     &              -0.7 &   -0.0 &             -1.1 &   3.5 &   2.5 &    0.0 \\
\hline
\textit{TMEM110}      &              -0.9 &    1.1 &             -1.0 &   3.3 &   3.1 &   -0.6 \\
\hline
\textit{IL17RA}       &              -4.4 &    3.1 &              7.9 &  -2.7 &  -2.5 &   -1.1 \\
\hline
\textit{IL18RAP}      &              -9.6 &   -0.9 &             -1.0 &  -1.6 &  -1.7 &    3.5 \\
\hline
\textit{ATP2A3}       &             -11.9 &   -0.8 &             -1.2 &  -2.1 &  -2.2 &    5.8 \\
\hline
\textit{FLG}          &             -20.1 &  -17.2 &             -6.0 &  -7.4 &  -8.5 &    4.0 \\
\hline
\end{tabular}
\label{asthma_table}
%\end{adjustwidth}
\end{table}

\subsubsection*{Glaucoma, intra-ocular pressure, and corneal resistance factor}
We also applied MRP to missense variants and PTVs for glaucoma, intra-ocular pressure, and corneal resistance factor as well as performing joint analyses. Intra-ocular pressure is a measure of the fluid pressure in the eye, is associated with glaucoma risk, and has been linked to genetic variants associated with glaucoma \cite{doi:10.1001/jama.2014.3192}. Corneal resistance factor is a measure of the cornea's ability to resist mechanical stress and has been associated with glaucoma presence and severity \cite{CXO:CXO414, MANSOURI2012419, GriseDulac:2012bo}. While the individual glaucoma analysis did not yield any associations with $\log_{10}$ BF greater than three, the joint analysis identified rare coding variants in both \textit{ANGPTL7}, \textit{WNT10A}, and \textit{KLHL22} as associated with glaucoma (Fig~\ref{glaucoma_bma}A-D). Applying the MRP model to identify protective effects also identified \textit{ANGPTL7} and \textit{WNT10A} but not \textit{KLHL22} (Fig~\ref{glaucoma_bma}E). We obtained similar results using the independent effects (\nameref{glaucoma_independent}) or similar effects models (\nameref{glaucoma_similar}).

Expression of \textit{ANGPTL7} is upregulated in glaucoma and has been proposed to regulate intra-ocular pressure and glaucoma risk \cite{Comes:2011ex, Kuchtey:2008ek}. \textit{KLHL22} has not been previously associated with glaucoma though a suggestive association was reported for retinopathy in individuals without diabetes \cite{10.1371/journal.pone.0054232}. The GWAS summary statistics for the rare variants in \textit{ANGPTL7} suggest that two missense variants, rs28991002 and rs28991009, appear to drive the association with glaucoma (Fig~\ref{glaucoma_bma}F).

\begin{figure}[!h]
\includegraphics[width=\textwidth]{../figures/final/Figure_04.pdf}
\caption{{\bf Results for glaucoma application.}
$\log_{10}$ Bayes Factors from applying MRP and Bayesian model averaging to summary statistics for missense and protein-truncating variants from (A) glaucoma (HC276), (B) intra-ocular pressure (INI5263), (C) corneal resistance factor (INI5265), and (D) all three traits jointly. (E) shows the results of a joint analysis focused on finding rare variants that protect against glaucoma. The genes outside of chromosome 6 with with Bayes Factor greater than three are indicated by arrows. Only $\log_{10}$ Bayes Factors greater than zero are plotted. F: $-\log_{10}$ $p$-values (left panel) and estimated effect sizes with 95\% confidence intervals (right panel) for missense variants and PTVs in \textit{ANGPTL7} for all three phenotypes.}
\label{glaucoma_bma}
\end{figure}

% %% TABLE: GLAUCOMA GENES OF INTEREST
\begin{table}[!ht]
\begin{adjustwidth}{-1.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf $\log_{10}$ Bayes Factors for genes highlighted in Figure~\ref{glaucoma_bma}.}}
\begin{tabular}{|l+p{3cm}|p{1cm}|p{1.6cm}|p{3.8cm}|p{3.4cm}|}
\hline
{\bf Gene} & {\bf Joint, protective} & {\bf Joint} & {\bf Glaucoma} & {\bf Intra-ocular pressure, Goldmann-correlated} & {\bf Corneal resistance factor} \\
\thickhline
\textit{ANGPTL7} &              10.1 &   9.4 &      1.7 &                                       10.1 &                       6.2 \\
\hline
\textit{KLHL22}  &               3.1 &   3.6 &     -0.2 &                                        3.9 &                       2.2 \\
\hline
\textit{GLI3}    &               2.9 &   2.9 &     -0.3 &                                       -0.7 &                       3.2 \\
\hline
\textit{WNT10A}  &               2.2 &   3.0 &     -0.3 &                                       -0.5 &                       3.5 \\
\hline
\textit{RNASE2}  &              -1.3 &  -0.3 &      0.4 &                                       -1.0 &                      -1.1 \\
\hline
\textit{FUK}     &              -6.5 &  -2.3 &      2.5 &                                       -2.1 &                      -2.1 \\
\hline
\end{tabular}
\label{table2}
\end{adjustwidth}
\end{table}

@@

\begin{figure}[H]
\includegraphics[width=\textwidth]{../figures/final/Figure_S01.pdf}
\end{figure}
\paragraph*{S1 Fig.}
\label{hdf5_fig}
{\bf HDF5 Implementation.} Our HDF5 implementation contained the following components: first, a group with one table per annotation file. All effect size (beta) values and study-specific annotations were contained here, and the number of tables is limited by $S\ \textrm{(the number of studies)} \times K\ \textrm{(the number of traits)}$. Second, a group with site-site covariance data. While these covariance matrices may have dimension $M\ \textrm{(the number of variants)} \times M$, we store the data as tables, each row specifiying the covariance between two variants. The number of tables should be the same as the previous set, capped by $S\ \textrm{(the number of studies)} \times K\ \textrm{(the number of traits)}$. Third, we store one table with sigma values for each study/phenotype combination. In the event that the traits were rank-normal transformation was performed these sigma values are equal to 1. These are used to compute correlation between two datasets. Finally, we store a matrix/table pair for Vy null and its index. The Vy null matrix has dimensions $ (S \times K) \times (S \times K)$ each entry specifying the estimated correlation of effect sizes between two datasets. The index table encodes row/column position of each dataset.

%% FIGURE: ASTHMA MANHATTAN AND FOREST FOR INDEPENDENT EFFECTS
\begin{figure}[H]
\includegraphics[width=\textwidth]{../figures/final/Figure_S02.pdf}
\end{figure}
\paragraph*{S2 Fig.}
\label{asthma_independent}
{\bf Results for independent effects model applied to asthma, eosinophil counts, FEV$_1$, and FVC.} $\log_{10}$ Bayes Factors from applying MRP independent effects model to summary statistics for missense and protein-truncating variants from (A) asthma (HC382), (B) eosinophil counts (INI30150), (C) forced vital capacity (FVC, INI3062), (D) forced expiratory volume in 1-second (FEV$_1$, INI3063), (E) all four traits jointly, and (F) all four traits jointly with focus on protective effects. The four genes outside of chromosome 6 with the largest Bayes Factors greater than three are labeled in each plot. Only $\log_{10}$ Bayes Factors greater than -5 are plotted.

%% FIGURE: ASTHMA MANHATTAN AND FOREST FOR SIMILAR EFFECTS
\begin{figure}[H]
\includegraphics[width=\textwidth]{../figures/final/Figure_S03.pdf}
\end{figure}
\paragraph*{S3 Fig.}
\label{asthma_similar}
{\bf Results for similar effects model applied to asthma, eosinophil counts, FEV$_1$, and FVC.} $\log_{10}$ Bayes Factors from applying MRP similar effects model to summary statistics for missense and protein-truncating variants from (A) asthma (HC382), (B) eosinophil counts (INI30150), (C) forced vital capacity (FVC, INI3062), (D) forced expiratory volume in 1-second (FEV$_1$, INI3063), (E) all four traits jointly, and (F) all four traits jointly with focus on protective effects. The four genes outside of chromosome 6 with the largest Bayes Factors greater than three are labeled in each plot. Only $\log_{10}$ Bayes Factors greater than -5 are plotted.

\begin{figure}[H]
\includegraphics[width=\textwidth]{../figures/final/Figure_S04.pdf}
\end{figure}
\paragraph*{S4 Fig.}
\label{glaucoma_independent}
{\bf Results for independent effects model applied to glaucoma intra-ocular pressure, and corneal resistance factor.} $\log_{10}$ Bayes Factors from applying MRP independent effects model to summary statistics for missense and protein-truncating variants from (A) glaucoma (HC276), (B) intra-ocular pressure (INI5263), (C) corneal resistance factor (INI5265), and (D) all three traits jointly. (E) shows the results of a joint analysis focused on finding rare variants that protect against glaucoma. The genes outside of chromosome 6 with with Bayes Factor greater than three are indicated by arrows. Only $\log_{10}$ Bayes Factors greater than zero are plotted.

\begin{figure}[H]
\includegraphics[width=\textwidth]{../figures/final/Figure_S05.pdf}
\end{figure}
\paragraph*{S5 Fig.}
\label{glaucoma_similar}
{\bf Results for similar effects model applied to glaucoma intra-ocular pressure, and corneal resistance factor.} $\log_{10}$ Bayes Factors from applying MRP similar effects model to summary statistics for missense and protein-truncating variants from (A) glaucoma (HC276), (B) intra-ocular pressure (INI5263), (C) corneal resistance factor (INI5265), and (D) all three traits jointly. (E) shows the results of a joint analysis focused on finding rare variants that protect against glaucoma. The genes outside of chromosome 6 with with Bayes Factor greater than three are indicated by arrows. Only $\log_{10}$ Bayes Factors greater than zero are plotted.
